Late and early effective start of non-pharmaceutical interventions inferred during COVID-19 outbreaks

### 2020-05-11 ms/ms.tex
table 2 read from tex

\usepackage{amssymb,amsmath,amsthm}
\usepackage{commath}
\usepackage{longtable}
\usepackage[hyphens]{url}
\usepackage[dvipsnames]{xcolor}
\usepackage[unicode=true,colorlinks=true,urlcolor=Blue,citecolor=black,linkcolor=black]{hyperref}
\def\equationautorefname~#1\null{Eq.~(#1)\null}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage{authblk}
\usepackage{lipsum}
\usepackage{multicol}
\usepackage{titlesec}	
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{enumitem}
\usepackage{soul}
\usepackage{booktabs}
\usepackage{pgfplotstable} % http://mirrors.ctan.org/graphics/pgf/contrib/pgfplots/doc/pgfplotstable.pdf
\usepackage{subcaption}

@@

%%% SEIR model %%%%
\paragraph*{SEIR model.}
We model SARS\=/CoV\=/2 infection dynamics by following the number of susceptible $S$, exposed $E$, reported infected $I_r$, and unreported infected $I_u$ individuals in a population of size $N$.
This model distinguishes between reported and unreported infected individuals: the reported infected are those that have enough symptoms to eventually be tested and thus appear in daily case reports, to which we fit the model.

Susceptible ($S$) individuals become exposed due to contact with reported or unreported infected individuals ($I_r$ or $I_u$) at a rate $\beta_t$ or $\mu \beta_t$.
The parameter $0 < \mu < 1$ represents the decreased transmission rate from unreported infected individuals, who are often subclinical or even asymptomatic.
The transmission rate $\beta_t \ge 0$ may change over  time $t$ due to behavioural changes of both susceptible and infected individuals.
Exposed individuals, after an average incubation period of $Z$ days, become reported infected with probability $\alpha_t$ or unreported infected with probability $(1-\alpha_t)$.
The reporting rate $0 < \alpha_t < 1$ may also change over time due to changes in human behavior.
Infected individuals remain infectious for an average period of $D$ days, after which they either recover, or becomes ill enough to be quarantined.
They therefore no longer infect other individuals, and the model does not track their frequency.
The model is described by the following equations:

@@

%%% Model fitting %%%
\paragraph*{Parameter estimation.}
To estimate the parameters of our model from the data $\vec{X}$, we apply a Bayesian inference approach.
We start our model $\Delta t$ days before the outbreak (defined as consecutive days with increasing confirmed cases) in each country~\citep{Gatto2020}.
The model in \autoref{eq:model} is parameterized by the vector $\theta$, where
\begin{equation} \label{eq:theta}
\theta=\Big(Z, D, \mu, \{\beta_t\}, \{\alpha_t\}, \{p_t\}, E(0), I_u(0)\Big), \tau, \Delta t.
\end{equation}

@@

%%% Fig 1 %%%
\begin{figure}[h]
    \centering
	\includegraphics[width=0.6\textwidth]{../figures/Fig1.pdf}
    \caption{
    \textbf{Official and effective start of non-pharmaceutical interventions.}
    	The difference between $\hat{\tau}$ the effective and $\tau^*$ the official start of NPI is shown for different regions. The effective NPI dates in Italy and Wuhan are significantly delayed compared to the official dates, whereas in Denmark, France, Spain, and Germany, the effective date is earlier than the official date.
	$\hat{\tau}$ is the posterior median, see \hl{TABLE2}. $\tau^*$ is the last NPI date, see \autoref{table:table1}. Thin and bold lines show 95\% and 75\% credible intervals (area in which $P(|\tau - \hat{\tau}| \mid \vec{X}) = 0.95$ and $0.75$.)
    }
    \label{fig:fig1}
\end{figure}

@@

@@

%%% Fig 4 %%%
\begin{figure}[h]
    \centering
	\includegraphics[width=0.5\textwidth]{../figures/Fig4.pdf}
    \caption{
    \textbf{Late effective start of NPIs leads to under-estimation of daily confirmed cases.}
    Real number of daily cases in Italy in black (markers: data, line: time moving average). Model predictions, assuming a 50\% decrease in transmission rate after the NPI starts, are shown as colored lines with 95\% confidence intervals. 
    Shaded box illustrates a serial interval of seven days.
    \textbf{(A)} Using the official date $\tau^*$ for the start of the NPI,  the model under-estimates the number of cases seven days after the start of the NPI.
    \textbf{(B)} Using the effective date $\hat{\tau}$ for the start of the NPI,  the model correctly estimates the number of cases seven days after the start of the NPI.
    Here, model parameters are MAP estimates for Italy (\hl{TABLE}) but with $\lambda=0.5$ and $\alpha_1=\alpha_2$.
    } 
    \label{fig:fig4}
\end{figure}

>>>>

\usepackage{amssymb,amsmath,amsthm}
\usepackage{commath}
\usepackage{longtable}
\usepackage[hyphens]{url}
\usepackage[dvipsnames]{xcolor}
\usepackage[unicode=true,colorlinks=true,urlcolor=Blue,citecolor=black,linkcolor=black]{hyperref}
\def\equationautorefname~#1\null{Eq.~(#1)\null}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage{authblk}
\usepackage{lipsum}
\usepackage{multicol}
\usepackage{titlesec}	
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{enumitem}
\usepackage{soul}
\usepackage{booktabs}
\usepackage{pgfplotstable} % http://mirrors.ctan.org/graphics/pgf/contrib/pgfplots/doc/pgfplotstable.pdf
\usepackage{subcaption}
\usepackage{dcolumn}
\newcolumntype{d}[1]{D{.}{.}{#1}}
\newcommand\ccell[1]{\multicolumn{1}{c}{#1}} % a wild guess...
\usepackage{adjustbox}

@@

%%% SEIR model %%%%
\paragraph*{SEIR model.} \label{sec:model}
We model SARS\=/CoV\=/2 infection dynamics by following the number of susceptible $S$, exposed $E$, reported infected $I_r$, and unreported infected $I_u$ individuals in a population of size $N$.
This model distinguishes between reported and unreported infected individuals: the reported infected are those that have enough symptoms to eventually be tested and thus appear in daily case reports, to which we fit the model.

Susceptible ($S$) individuals become exposed due to contact with reported or unreported infected individuals ($I_r$ or $I_u$) at a rate $\beta_t$ or $\mu \beta_t$.
The parameter $0 < \mu < 1$ represents the decreased transmission rate from unreported infected individuals, who are often subclinical or even asymptomatic.
The transmission rate $\beta_t \ge 0$ may change over  time $t$ due to behavioral changes of both susceptible and infected individuals.
Exposed individuals, after an average incubation period of $Z$ days, become reported infected with probability $\alpha_t$ or unreported infected with probability $(1-\alpha_t)$.
The reporting rate $0 < \alpha_t < 1$ may also change over time due to changes in human behavior.
Infected individuals remain infectious for an average period of $D$ days, after which they either recover, or becomes ill enough to be quarantined.
They therefore no longer infect other individuals, and the model does not track their frequency.
The model is described by the following equations:

@@

%%% Model fitting %%%
\paragraph*{Parameter estimation.}
To estimate the parameters of our model from the data $\vec{X}$, we apply a Bayesian inference approach.
We start our model $\Delta t$ days before the outbreak (defined as consecutive days with increasing confirmed cases) in each country~\citep{Gatto2020}.
The model in \autoref{eq:model} is parameterized by the vector $\theta$, where
\begin{equation} \label{eq:theta}
\theta=\Big(Z, D, \mu, \{\beta_t\}, \{\alpha_t\}, \{p_t\}, E(0), I_u(0), \tau, \Delta t \Big).
\end{equation}

@@

%%% Fig 1 %%%
\begin{figure}[h]
    \centering
	\includegraphics[width=0.6\textwidth]{../figures/Fig1.pdf}
    \caption{
    \textbf{Official and effective start of non-pharmaceutical interventions.}
    	The difference between $\hat{\tau}$ the effective and $\tau^*$ the official start of NPI is shown for different regions. The effective NPI dates in Italy and Wuhan are significantly delayed compared to the official dates, whereas in Denmark, France, Spain, and Germany, the effective date is earlier than the official date.
	$\hat{\tau}$ is the posterior median, see \autoref{table:table2}. $\tau^*$ is the last NPI date, see \autoref{table:table1}. Thin and bold lines show 95\% and 75\% credible intervals (area in which $P(|\tau - \hat{\tau}| \mid \vec{X}) = 0.95$ and $0.75$.)
    }
    \label{fig:fig1}
\end{figure}

@@

%% Table 2 %%%
\begin{table}[h]
\centering
\scriptsize
\input{../figures/Table2.tex}
\caption{
\textbf{Parameter estimates for different regions.}
See \autoref{eq:model} for model parameters.
All estimates are posterior medians.
75\% and 95\% credible intervals given only for $\tau$, in days.
$\tau^*$ is the official last NPI date, see \autoref{table:table1}.
}
\label{table:table2}
\end{table}

@@

%%% Fig 4 %%%
\begin{figure}[h]
    \centering
	\includegraphics[width=0.5\textwidth]{../figures/Fig4.pdf}
    \caption{
    \textbf{Late effective start of NPIs leads to under-estimation of daily confirmed cases.}
    Real number of daily cases in Italy in black (markers: data, line: time moving average). Model predictions, assuming a 50\% decrease in transmission rate after the NPI starts, are shown as colored lines with 95\% confidence intervals. 
    Shaded box illustrates a serial interval of seven days.
    \textbf{(A)} Using the official date $\tau^*$ for the start of the NPI,  the model under-estimates the number of cases seven days after the start of the NPI.
    \textbf{(B)} Using the effective date $\hat{\tau}$ for the start of the NPI,  the model correctly estimates the number of cases seven days after the start of the NPI.
    Here, model parameters are estimates for Italy (\autoref{table:table2}) but with $\lambda=0.5$ and $\alpha_1=\alpha_2$.
    } 
    \label{fig:fig4}
\end{figure}
